---
title: "TP: Environmental Risk"
subtitle: "Work Sheet 2"
format: 
  live-html:
    toc: true
    toc-location: left
    theme: [simple, "presentations/styles/style.scss"]
webr:
  packages:
    - tidyverse
  render-df: gt-interactive
live:
  show-hints: true
  show-solutions: true    
highlight-style: arrow
---

{{< include ./_extensions/r-wasm/live/_knitr.qmd >}}

# Task 2.1: Explore the `coral_cover` data set

Data is already loaded:

```{r}
#| echo: true
#| eval: false
# Don't forget to load the tidyverse package:
library(tidyverse)

# Read the coral_cover data:
dat_coral_cover <- read.csv("https://raw.githubusercontent.com/andieich/practicals_envrisk/refs/heads/main/data/coral_cover.csv")
```

```{r}
#| label: tbl-coral_cover
#| tbl-cap: First five rows of `coral_cover` data set
#| echo: false
#| warnings: false
#| message: false

library(gt)
library(gtExtras)
library(tidyverse)

dat_coral_cover <- read.csv("https://raw.githubusercontent.com/andieich/practicals_envrisk/refs/heads/main/data/coral_cover.csv") 

dat_coral_cover %>% 
  head(5) %>% 
  gt()  %>% 
  gt_theme_excel() %>% 
  tab_options(table.align = "center", 
              table.font.size = pct(80))
```

You can use R functions to answer the questions or [open the data frame in Excel:](https://raw.githubusercontent.com/andieich/practicals_envrisk/refs/heads/main/data/coral_cover.xlsx)

::: med-space
:::

### Which years are covered?

::: med-space
:::

::::: panel-tabset
## Task

```{webr}
#| exercise: 
#|   - ex_2.1a
#|   - ex_2.1b
#|   - ex_2.1c
#|   - ex_2.1d
#|   - ex_2.2
#|   - ex_2.3
#| setup: true

library(tidyverse)

# Read the coral_cover data:
dat_coral_cover <- read.csv("https://raw.githubusercontent.com/andieich/practicals_envrisk/refs/heads/main/data/coral_cover.csv")
```

```{webr}
#| exercise: ex_2.1a
#| warnings: false
#| message: false
dat_coral_cover %>% 
  ______
```

## Hints

::: {.hint exercise="ex_2.1a"}
``` r
dat_coral_cover %>%    #<1>
  summarise(______ = ______,   #<2>
            ______ = ______)   #<3>
```

1.  Take the `dat_coral_cover` data set, and then,
2.  calculate the `min()`
3.  and `max()` of `year`
:::

## Solution

::: {.solution exercise="ex_2.1a"}
``` r
dat_coral_cover %>%   #<1>
  summarise(min = min(year),  #<2>
            max = max(year))    #<3>
```

1.  take the `dat_coral_cover` data set, and then,
2.  calculate the `min()`
3.  and `max()` of `year`

*Solution: 2005 to 2024*
:::
:::::

### Are there any gaps?

::: med-space
:::

::::: panel-tabset
## Task

```{webr}
#| exercise: ex_2.1b
#| warnings: false
#| message: false
dat_coral_cover %>% 
  ______
```

## Hints

::: {.hint exercise="ex_2.1b"}
``` r
dat_coral_cover %>%     #<1>
  select(______) %>%    #<2> 
  distinct() %>%     #<3>
  summarise(______ = diff(______))     %>% #<4>
  filter(______ != ______)     #<5>
```

1.  Take the `dat_coral_cover` data set, and then,
2.  only select the `year` column, then,
3.  remove any duplicated rows with `disticnt()`, then,
4.  calculate difference between years with `diff(columnname)`, then,
5.  show only rows in which the difference is not (`!=`) 1
:::

## Solution

::: {.solution exercise="ex_2.1b"}
``` r
dat_coral_cover %>%     #<1>
  select(year) %>%    #<2> 
  distinct() %>%     #<3>
  summarise(diff_year = diff(year))     %>% #<4>
  filter(diff_year != 1)     #<5>
```

1.  Take the `dat_coral_cover` data set, and then,
2.  only select the `year` column, then,
3.  remove any duplicated rows, then,
4.  calculate difference between years, then,
5.  show only rows in which the difference is not 1

*Solution: No gaps*
:::
:::::

### How many sites?

::: med-space
:::

::::: panel-tabset
## Task

```{webr}
#| exercise: ex_2.1c
#| warnings: false
#| message: false
dat_coral_cover %>% 
  ______
```

## Hints

::: {.hint exercise="ex_2.1c"}
``` r
dat_coral_cover %>%      #<1>
  select(______) %>%      #<2>
  distinct() %>%      #<3>
  summarise(______ = ______)     #<4>
```

1.  Take the `dat_coral_cover` data set, and then,
2.  only select the `site` column, then,
3.  remove any duplicated rows, then,
4.  show only the number of rows with `n()`
:::

## Solution

::: {.solution exercise="ex_2.1c"}
``` r
dat_coral_cover %>%      #<1>
  select(site) %>%      #<2>
  distinct() %>%      #<3>
  summarise(n = n())     #<4>
```

1.  Take the `dat_coral_cover` data set, and then,
2.  only select the `site` column, then,
3.  remove any duplicated rows, then,
4.  show only the number of rows

*Solution: 13 sites*
:::
:::::

### Same number of transects per site and year?

::: med-space
:::

::::: panel-tabset
## Task

```{webr}
#| exercise: ex_2.1d
#| warnings: false
#| message: false
dat_coral_cover %>% 
  ______
```

## Hints

::: {.hint exercise="ex_2.1d"}
``` r
dat_coral_cover %>%       #<1>
  group_by(______, ______) %>%       #<2>
  summarise(______ = ______ %>%       #<3>
  filter(______ != ______)      #<4>
```

1.  Take the `dat_coral_cover` data set, and then,
2.  group by `site` and `year`, then,
3.  calculate the number or rows per group with `n()`, then
4.  show only rows in the number of rows is not (`!=`) 3
:::

## Solution

::: {.solution exercise="ex_2.1d"}
``` r
dat_coral_cover %>%       #<1>
  group_by(site, year) %>%       #<2>
  summarise(n = n()) %>%       #<3>
  filter(n != 3)      #<4>
```

1.  Take the `dat_coral_cover` data set, and then,
2.  group by `site` and `year`, then,
3.  calculate the number or rows per group, then
4.  show only rows in the number of rows is not 3

*Solution: Always 3 transects per site and year*
:::
:::::

# Task 2.2: Make a similar plot

<br>

```{r}
#| echo: false
#| message: false


dat_coral_cover_sitesS <- dat_coral_cover %>% 
  group_by(year, site) %>% 
  summarise(percent = mean(percent)) 


dat_coral_coverS <- dat_coral_cover_sitesS %>% 
  group_by(year) %>% 
  summarise(mean_percent = mean(percent),
            sd = sd(percent))

plot_cover <- dat_coral_cover_sitesS %>% 
  ggplot(aes(x = year))+
  geom_ribbon(data = dat_coral_coverS, 
              aes(ymin = mean_percent - sd, ymax = mean_percent + sd), fill = "grey80")+
    geom_line(aes(y = mean_percent),
            data = dat_coral_coverS, linewidth = 1)+
  geom_point(aes(y = percent),
             position = position_jitter(width = .2, seed = 1),
             shape = 21, fill = "darkgreen", col = "white", size = 3)+

  labs(x = NULL, y = "Coral Cover (%)")
plot_cover+
  theme_minimal(base_size = 18)
```

-   The points are the mean `percent` values per site
-   The line is the mean of these mean values
-   The shaded area in this mean ± the standard deviation `sd()`

::: med-space
:::

::::: panel-tabset
## Task

```{webr}
#| exercise: ex_2.2
#| warnings: false
#| message: false

# Calculate the mean coral cover per site and year
dat_coral_cover_sitesS <- dat_coral_cover %>% 
  ______

# For each year, calculate the mean and sd of the mean cover per site
dat_coral_coverS <- dat_coral_cover_sitesS %>% 
  ______

# Plot the data
dat_coral_cover_sitesS %>% 
  ggplot()+
  
  #  a ribbon layer
  geom_ribbon(data = ______, 
              aes(ymin = ______, ymax = ______), fill = "grey80")+
  
  # add a line
  geom_line(data = ______,
            aes(y = ______))+
  
  # add the raw data
  geom_point(aes(y = ______),
             position = position_jitter(width = .2))
```

## Hints

::: {.hint exercise="ex_2.2"}
``` r
# Calculate the mean coral cover per site and year
dat_coral_cover_sitesS <- dat_coral_cover %>%       #<1>
  group_by(______, ______) %>%       #<2>
  summarise(percent = mean(______))       #<3>

# For each year, calculate the mean and sd of the mean cover per site
dat_coral_coverS <- dat_coral_cover_sitesS %>%       #<4>
  group_by(______) %>%       #<5>
  summarise(mean_percent = mean(______),      #<6>
            sd = sd(______))      #<7>
            
# Plot the data
  ggplot(mapping = aes(x = ______))+      #<8>
  geom_ribbon(data = ______,       #<9>
              aes(ymin = ______, ymax = ______), fill = "grey80")+      #<10>
  geom_line(data = ______, #<11>
            aes(y = ______)) +    #<12>
  geom_point(data = ______, #<13>
             aes(y = ______),      #<14>
             position = position_jitter(width = .2, seed = 1))  +    #<15>
  labs(x = "______", y = "______") +     #<16>
  theme_minimal()  #<17>
```

1.  Make a new summary data frame called `dat_coral_cover_sitesS`. Take the `dat_coral_cover` data frame and
2.  group it by `year` and `site`, then,
3.  calculate the mean percent value.
4.  Then, create another summary data frame called `dat_coral_coverS`. Take the just created `dat_coral_cover_sitesS` summary data frame and
5.  group it by `year`, then,
6.  calculate the mean
7.  and standard deviation (`sd()`). Now you are ready to plot the data.
8.  Make a new ggplot, store in `aes()` the axes that will be used in all layers. `year` should be plotted on the x axes, the variable names for the y axis differs for the layers (`percent` and `mean_percent`).
9.  Add a `geom_ribbon()`. Use `dat_coral_coverS` as input data
10. and `mean_percent - sd` for `ymin` and `mean_percent - sd` for `ymax`. You can define the fill colour with `fill`.
11. Add a `geom_line()`. Use `dat_coral_coverS` as input data
12. and `mean_percent` as y axis. Then,
13. add a `geom_point()`. This time, use `dat_coral_cover_sitesS` as input data
14. and `percent` as y axis.
15. You can randomly wiggle the position of the points to show the variability with `position_jitter()`.
16. Define the axis titles
17. and select a theme.
:::

## Solution

::: {.solution exercise="ex_2.2"}
``` r
# Calculate the mean coral cover per site and year
dat_coral_cover_sitesS <- dat_coral_cover %>%       #<1>
  group_by(year, site) %>%       #<2>
  summarise(percent = mean(percent))       #<3>

# For each year, calculate the mean and sd of the mean cover per site
dat_coral_coverS <- dat_coral_cover_sitesS %>%       #<4>
  group_by(year) %>%       #<5>
  summarise(mean_percent = mean(percent),      #<6>
            sd = sd(percent))      #<7>
            
# Plot the data
  ggplot(mapping = aes(x = year))+      #<8>
  geom_ribbon(data = dat_coral_coverS,       #<9>
              aes(ymin = mean_percent - sd, ymax = mean_percent + sd), fill = "grey80")+      #<10>
  geom_line(data = dat_coral_coverS, #<11>
            aes(y = mean_percent)) +     #<12>
  geom_point(data = dat_coral_cover_sitesS, #<13>
             aes(y = percent),      #<14>
             position = position_jitter(width = .2, seed = 1)) +    #<15>
  labs(x = NULL, y = "Coral Cover (%)") +     #<16>
  theme_minimal()  #<17>
```

1.  Make a new summary data frame called `dat_coral_cover_sitesS`. Take the `dat_coral_cover` data frame and
2.  group it by `year` and `site`, then,
3.  calculate the mean percent value.
4.  Then, create another summary data frame called `dat_coral_coverS`. Take the just created `dat_coral_cover_sitesS` summary data frame and
5.  group it by `year`, then,
6.  calculate the mean
7.  and standard deviation (`sd()`). Now you are ready to plot the data.
8.  Make a new ggplot, store in `aes()` the axes that will be used in all layers. `year` should be plotted on the x axes, the variable names for the y axis differs for the layers (`percent` and `mean_percent`).
9.  Add a `geom_ribbon()`. Use `dat_coral_coverS` as input data
10. and `mean_percent - sd` for `ymin` and `mean_percent - sd` for `ymax`. You can define the fill colour with `fill`.
11. Add a `geom_line()`. Use `dat_coral_coverS` as input data
12. and `mean_percent` as y axis. Then,
13. add a `geom_point()`. This time, use `dat_coral_cover_sitesS` as input data
14. and `percent` as y axis.
15. You can randomly wiggle the position of the points to show the variability with `position_jitter()`.
16. Define the axis titles
17. and select a theme.
:::
:::::



# Task 2.3: Explore variability between sites

1.  Filter the `dat_coral_cover` for `year >= 2019` and `year <= 2023`

2.  Make a summary data frame with the mean per `site` and `year`.


3.  Find way to plot the variability between sites. For example, you could plot separate lines for each site or make sub-panels with `facet_wrap()`.


::: med-space
:::

::::: panel-tabset
## Task



```{webr}
#| exercise: ex_2.3
#| warnings: false
#| message: false

# Filter dat_coral_cover for `year >= 2019` and `year <= 2023`
dat_coral_cover_2019 <- dat_coral_cover %>% 
  ______

# For each site and year, calculate the mean cover
dat_coral_cover_2019_S <- dat_coral_cover_2019 %>% 
  ______

# Version 1: Plot lines on top of each other
ggplot(mapping = aes(x = ______,
                       col = ______))+

  # add a line
  geom_line(data = ______,
            aes(y = ______))+
  
  # add the raw data
  geom_point(aes(y = ______),
             position = position_jitter(width = .2))
  
  
# Version 2: Different sub-panels for sites
ggplot(mapping = aes(x = ______,
                       col = ______))+

  # add a line
  geom_line(data = ______,
            aes(y = ______))+
  
  # add the raw data
  geom_point(aes(y = ______),
             position = position_jitter(width = .2)) + 
    
  # Separate into sub-panels
  facet_wrap(~______)
```

## Hints

::: {.hint exercise="ex_2.3"}

``` r
# Filter dat_coral_cover for `year >= 2019` and `year <= 2023`
dat_coral_cover_2019 <- ______ %>%       #<1>
  filter(______ >= ______, ______ <= ______)      #<2>
  

# For each site and year, calculate the mean cover
dat_coral_cover_2019_S <- ______ %>%       #<3>
  group_by(______, ______) %>%       #<4>
  summarise(mean_percent = ______(______))      #<5>

# Version 1: Plot lines on top of each other
ggplot(mapping = aes(x = ______,      #<6>
                       col = ______))+      #<7>

  # add a line
  geom_line(data = ______,      #<8>
            aes(y = ______))+      #<9>
  
  # add the raw data
  geom_point(data = ______,      #<10>
             aes(y = ______),      #<11>
             position = position_jitter(width = .2))+      #<12>
  labs(x = "______", y = "______")+                 #<13>
  theme_minimal()+      #<14>
  theme(legend.position = "None")      #<14>
  
  
# Version 2: Different sub-panels for sites
ggplot(mapping = aes(x = ______,     
                       col = ______))+  

  # add a line
  geom_line(data = ______,  
            aes(y = ______))+  
  
  # add the raw data
  geom_point(data = ______,  
             aes(y = ______),
             position = position_jitter(width = .2))+  
  labs(x = "______", y = "______")+
  facet_wrap(~______) + #<15>
  theme_minimal()+    
  theme(legend.position = "None")   
  
```

1. Make a new data frame for the filtered data called `dat_coral_cover_2019`. Take the `dat_coral_cover` data frame and 
2. filter for `year >= 2019` and `year <= 2023`.
3. Make a new summary data frame for the filtered data called `dat_coral_cover_2019_S`. Take the filtered `dat_coral_cover_2019` data, 
4. group by `site` and `year`,
5. and calculate the mean cover.
6. Then, plot the data.  Store in `aes()` the axes that will be used in all layers. `year` should be plotted on the x axes.
7. `site` should be used to color the points and lines. If used like that, separate lines will be drawn for each site.
8. Add a `geom_line()` for the `dat_coral_cover_2019_S` data to plot
9. the `mean_percent` data.
10. Add a `geom_point()` for the `dat_coral_cover_2019` data 
11. to plot the `percent` data
12. You can randomly wiggle the position of the points to show the variability with `position_jitter()`.
13. Define the axis titles
14. and select a theme.
15. To separate the sites into subplot, keep everything as before and add `facet_wrap(~column_name)` for `site`.
:::

## Solution

::: {.solution exercise="ex_2.3"}

``` r
# Filter dat_coral_cover for `year >= 2019` and `year <= 2023`
dat_coral_cover_2019 <- dat_coral_cover %>%       #<1>
  filter(year >= 2019, year <= 2023)      #<2>
  

# For each site and year, calculate the mean cover
dat_coral_cover_2019_S <- dat_coral_cover_2019 %>%       #<3>
  group_by(site, year) %>%       #<4>
  summarise(mean_percent = mean(percent))      #<5>

# Version 1: Plot lines on top of each other
ggplot(mapping = aes(x = year,      #<6>
                       col = site))+      #<7>

  # add a line
  geom_line(data = dat_coral_cover_2019_S,      #<8>
            aes(y = mean_percent))+      #<9>
  
  # add the raw data
  geom_point(data = dat_coral_cover_2019,      #<10>
             aes(y = percent),      #<11>
             position = position_jitter(width = .2))+      #<12>
  labs(x = NULL, y = "Coral Cover (%)")+                 #<13>
  theme_minimal()+      #<14>
  theme(legend.position = "None")      #<14>
  
  
# Version 2: Different sub-panels for sites
ggplot(mapping = aes(x = year,     
                       col = site))+  

  # add a line
  geom_line(data = dat_coral_cover_2019_S,  
            aes(y = mean_percent))+  
  
  # add the raw data
  geom_point(data = dat_coral_cover_2019,  
             aes(y = percent),
             position = position_jitter(width = .2))+  
  labs(x = NULL, y = "Coral Cover (%)")+
  facet_wrap(~site) + #<15>
  theme_minimal()+    
  theme(legend.position = "None")   
  
```

1. Make a new data frame for the filtered data called `dat_coral_cover_2019`. Take the `dat_coral_cover` data frame and 
2. filter for `year >= 2019` and `year <= 2023`.
3. Make a new summary data frame for the filtered data called `dat_coral_cover_2019_S`. Take the filtered `dat_coral_cover_2019` data, 
4. group by `site` and `year`,
5. and calculate the mean cover.
6. Then, plot the data.  Store in `aes()` the axes that will be used in all layers. `year` should be plotted on the x axes.
7. `site` should be used to color the points and lines. If used like that, separate lines will be drawn for each site.
8. Add a `geom_line()` for the `dat_coral_cover_2019_S` data to plot
9. the `mean_percent` data.
10. Add a `geom_point()` for the `dat_coral_cover_2019` data 
11. to plot the `percent` data
12. You can randomly wiggle the position of the points to show the variability with `position_jitter()`.
13. Define the axis titles
14. and select a theme.
15. To separate the sites into subplot, keep everything as before and add `facet_wrap(~______)` for `site`.
:::
:::::


# Task 2.4: Combine data


Combine `dat_change_coral_cover` and `dat_max_dhw_2019_sites` based on the `site` column

Use the `left_join()` function:

```{r}
#| eval: false
#| echo: true

dat_new <- left_join(x = data_1, #<1>
                     y = data_1, #<2>
                     by = "column_name") #<3>
                      
```

1. `data_1` will be used as "template". To all rows data will be added. If not in `data_2`, `NA` values will be added
2. `data_2` will be used to add all columns to `data_1`, 
3. based on common values in the `"column_name"` (Note the `""`)




::: med-space
:::

::::: panel-tabset
## Task

```{webr}
#| exercise: 
#|   - ex_2.4
#| setup: true

library(tidyverse)

# Read the coral_cover data:
dat_coral_cover <- read.csv("https://raw.githubusercontent.com/andieich/practicals_envrisk/refs/heads/main/data/coral_cover.csv")

dat_max_dhw_2019_sites <- read.csv("https://raw.githubusercontent.com/andieich/practicals_envrisk/refs/heads/main/data/max_dhw_2019_sites.csv")

dat_change_coral_cover <- dat_coral_cover %>% 
  filter(year >= 2019, year <= 2020) %>% 
  group_by(site, year) %>% 
  summarise(percent = mean(percent)) %>% 
  pivot_wider(names_from = year, 
              values_from = percent) %>% 
  # calculate relative change
  mutate(rel_change = 100*(`2019` - `2020`)/`2019`) %>% 
  select(-`2020`, -`2019`) %>% 
  ungroup()

```

```{webr}
#| exercise: ex_2.4
#| warnings: false
#| message: false
# combine data
dat_change_coral_cover <- left_join(______, 
                                    ______,  
                                    by = "______") 
# show combined data
dat_change_coral_cover
```



## Hints

::: {.hint exercise="ex_2.4"}

``` r
dat_change_coral_cover <- left_join(______,  #<1>
                                    ______, #<2>
                                    by = "______") #<3>
```
1. Replace `dat_change_coral_cover` with the combined data of `dat_change_coral_cover` and `dat_max_dhw_2019_sites`. Use `dat_change_coral_cover` as a template 
2. to add data from `dat_max_dhw_2019_sites` 
3. based on common values in the `"sites"` column

:::

## Solution

::: {.solution exercise="ex_2.4"}

``` r
dat_change_coral_cover <- left_join(dat_change_coral_cover,  #<1>
                                    dat_max_dhw_2019_sites,  #<2>
                                    by = "site")  #<3>
```
1. Replace `dat_change_coral_cover` with the combined data of `dat_change_coral_cover` and `dat_max_dhw_2019_sites`. Use `dat_change_coral_cover` as a template 
2. to add data from `dat_max_dhw_2019_sites` 
3. based on common values in the `"sites"` column

:::
:::::


# Task 2.5: Regression


Analyse the impact of `max_dhw` on `rel_change`

-   Is this effect *significant*?

-   What is the expected change in coral cover per unit DHW?



::: med-space
:::

::::: panel-tabset
## Task

```{webr}
#| exercise: 
#|   - ex_2.5
#|   - ex_2.6
#| setup: true

library(tidyverse)

# Read the coral_cover data:
dat_coral_cover <- read.csv("https://raw.githubusercontent.com/andieich/practicals_envrisk/refs/heads/main/data/coral_cover.csv")

dat_max_dhw_2019_sites <- read.csv("https://raw.githubusercontent.com/andieich/practicals_envrisk/refs/heads/main/data/max_dhw_2019_sites.csv")

dat_change_coral_cover <- dat_coral_cover %>% 
  filter(year >= 2019, year <= 2020) %>% 
  group_by(site, year) %>% 
  summarise(percent = mean(percent)) %>% 
  pivot_wider(names_from = year, 
              values_from = percent) %>% 
  # calculate relative change
  mutate(rel_change = 100*(`2019` - `2020`)/`2019`) %>% 
  select(-`2020`, -`2019`) %>% 
  ungroup()

dat_change_coral_cover <- left_join(dat_change_coral_cover,
                                    dat_max_dhw_2019_sites, 
                                    by = "site")

# for ex_2.6
m_change_coral_cover <- lm(rel_change ~ max_dhw,  
                           data = dat_change_coral_cover) 
```

```{webr}
#| exercise: ex_2.5
#| warnings: false
#| message: false
# Run a regression
m_change_coral_cover <- lm(______ ~ ______, 
                           data = ______)

# Show results
summary(______) 
```



## Hints

::: {.hint exercise="ex_2.5"}

``` r
# Run a regression
m_change_coral_cover <- lm(______ ~ ______,    #<1>
                           data = ______)   #<2>

# Show results
summary(______)  #<3>
```
1. Run a regression model and call it `m_change_coral_cover`. Use `max_dhw` to explain `rel_change`.
2. Both columns are in the `dat_change_coral_cover` data frame.
3. Show the results of the regression called `m_change_coral_cover` 
:::

## Solution

::: {.solution exercise="ex_2.5"}

``` r
# Run a regression
m_change_coral_cover <- lm(rel_change ~ max_dhw,    #<1>
                           data = dat_change_coral_cover)   #<2>

# Show results
summary(m_change_coral_cover)  #<3>
```

1. Run a regression model and call it `m_change_coral_cover`. Use `max_dhw` to explain `rel_change`.
2. Both columns are in the `dat_change_coral_cover` data frame.
3. Show the results of the regression called `m_change_coral_cover` 


::: justify
> ### Meaning
>
> The intensity of the heat stress, measured as the maximum degree heating weeks in 2019 (DHW~max~), significantly impacted the decline in coral cover (t(11) = 5.036, p \< .001, adj. R^2^ = 0.67). The expected relative decline in coral cover increased by 90.5% ± 18.0% (SE) per unit of DHW~max~.
:::

:::
:::::

# Task 2.6: Regression visualisation


Plot the coral cover data and model in a similar way

```{r}
#| echo: false
#| message: false

# Read the coral_cover data:
dat_coral_cover <- read.csv("https://raw.githubusercontent.com/andieich/practicals_envrisk/refs/heads/main/data/coral_cover.csv")

dat_max_dhw_2019_sites <- read.csv("https://raw.githubusercontent.com/andieich/practicals_envrisk/refs/heads/main/data/max_dhw_2019_sites.csv")

dat_change_coral_cover <- dat_coral_cover %>% 
  filter(year >= 2019, year <= 2020) %>% 
  group_by(site, year) %>% 
  summarise(percent = mean(percent)) %>% 
  pivot_wider(names_from = year, 
              values_from = percent) %>% 
  # calculate relative change
  mutate(rel_change = 100*(`2019` - `2020`)/`2019`) %>% 
  select(-`2020`, -`2019`) %>% 
  ungroup()


dat_change_coral_cover <- left_join(dat_change_coral_cover,
                                    dat_max_dhw_2019_sites, 
                                    by = "site")

m_change_coral_cover <- lm(rel_change ~ max_dhw,
                       data = dat_change_coral_cover)


ndat_coralcover_dhw <-  data.frame(max_dhw = seq(min(dat_change_coral_cover$max_dhw),
                                        max(dat_change_coral_cover$max_dhw),
                                        length = 100))

pred_coralcover_dhw  <- predict(m_change_coral_cover,
                       newdata = ndat_coralcover_dhw,
                       se.fit = T) %>% 
  bind_cols(ndat_coralcover_dhw)
  

pred_coralcover_dhw %>% 
  
  ggplot(aes(x = max_dhw))+
  
  # Plot SE
  geom_ribbon(aes(ymin = fit - se.fit,
                  ymax = fit + se.fit),
              fill = "grey")+
  
  #Plot model
  geom_line(aes(y = fit), 
            col = "darkred", linewidth = 1.2)+
  
  # Plot raw data
  geom_point(data = dat_change_coral_cover,
             aes(y = rel_change), size = 4)+
  
  # Formatting
  labs(x = expression(DWH[max]),
       y = expression(Delta~Relative~cover~"(%)"))+
  theme_light()+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
```




::: med-space
:::

::::: panel-tabset
## Task



```{webr}
#| exercise: ex_2.6
#| warnings: false
#| message: false
# Make a new data frame with sequence of x values
ndat_coralcover_dhw <-  data.frame(______ = seq(min(______),
                                                max(______),                                                                 length = 100))

# Predict data
pred_coralcover_dhw  <- predict(______,
                                newdata = ______,
                                se.fit = T) %>% 
  bind_cols(______)


# Plot
ggplot(mapping = aes(x = ______))+
  # Plot SE
  geom_ribbon(aes(ymin = ______,
                  ymax = ______),
              fill = "grey")+
  #Plot model
  geom_line(aes(y = ______), 
            col = "darkred", linewidth = 1.2)+
  # Plot raw data
  geom_point(data = ______,
             aes(y = ______),
             size = 4)


```



## Hints

::: {.hint exercise="ex_2.6"}

``` r
# Make a new data frame with sequence of x values
ndat_coralcover_dhw <-  data.frame(______ = seq(min(______),     #<1>
                                                 max(______,    #<2>
                                                 length = ______))    #<3>

# Predict data
pred_coralcover_dhw  <- predict(______,    #<4>
                                newdata = ______,    #<5>
                                se.fit = T) %>%     #<6>
                        bind_cols(______)    #<7>
  
# Plot
ggplot(mapping = aes(x = ______))+    #<8>
  # Plot SE
  geom_ribbon(data = ______,    #<9>
              aes(ymin = ______,    #<9>
                  ymax = ______),    #<9>
              fill = "grey")+
  #Plot model
  geom_line(data = ______,    #<10>
            aes(y = ______),     #<10>
            col = "darkred", linewidth = 1.2)+   
  # Plot raw data
  geom_point(data = ______,    #<11>
             aes(y = ______), size = 4)    #<11>
```
1. Make a new data frame and call it `ndat_coralcover_dhw`. Add a column called `max_dhw` containing a sequence from the minimum  `max_dhw` value in `dat_change_coral_cover` ()
2. To the maximum `max_dhw` value.
3. The sequence should have 100 entries
4. Based on the model `m_change_coral_cover`, predict values and store them in a new data frame called `pred_coralcover_dhw`.
5. The predicted values should be done for `x` values in `ndat_coralcover_dhw`.
6. Also predict the standard error.
7. `predict()` only returns the predicted values. For plotting, add again the `x` values stored in `ndat_coralcover_dhw`. Now you are set up for the plot.
8. Store in `aes()` the axes that will be used in all layers. `max_dhw` should be plotted on the x axes.
9. Add a `geom_ribbon()` for the predicted data ± the standard error (`fit - se.fit` for `ymin`and `fit + se.fit` for `ymax`). These columns are in `pred_coralcover_dhw`.
10. Now, add the regression line with `geom_line()`. The column `fit` in `pred_coralcover_dhw` should be used as y axis.
11. Then, add the raw data that is stored in `dat_change_coral_cover`. Use `rel_change` as y axis.




:::

## Solution

::: {.solution exercise="ex_2.6"}

``` r
# Make a new data frame with sequence of x values
ndat_coralcover_dhw <-  data.frame(max_dhw = seq(min(dat_change_coral_cover$max_dhw),     #<1>
                                                 max(dat_change_coral_cover$max_dhw),    #<2>
                                                 length = 100))    #<3>

# Predict data
pred_coralcover_dhw  <- predict(m_change_coral_cover,    #<4>
                                newdata = ndat_coralcover_dhw,    #<5>
                                se.fit = T) %>%     #<6>
                        bind_cols(ndat_coralcover_dhw)    #<7>
  
# Plot
ggplot(mapping = aes(x = max_dhw))+    #<8>
  # Plot SE
  geom_ribbon(data = pred_coralcover_dhw,    #<9>
              aes(ymin = fit - se.fit,    #<9>
                  ymax = fit + se.fit),    #<9>
              fill = "grey")+
  #Plot model
  geom_line(data = pred_coralcover_dhw,    #<10>
            aes(y = fit),     #<10>
            col = "darkred", linewidth = 1.2)+   
  # Plot raw data
  geom_point(data = dat_change_coral_cover,    #<11>
             aes(y = rel_change), size = 4) +   #<11>
  # Formatting
  labs(x = expression(DWH[max]),
       y = expression(Delta~Relative~cover~"(%)"))+
  theme_light()+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
```
1. Make a new data frame and call it `ndat_coralcover_dhw`. Add a column called `max_dhw` containing a sequence from the minimum  `max_dhw` value in `dat_change_coral_cover` ()
2. To the maximum `max_dhw` value.
3. The sequence should have 100 entries
4. Based on the model `m_change_coral_cover`, predict values and store them in a new data frame called `pred_coralcover_dhw`.
5. The predicted values should be done for `x` values in `ndat_coralcover_dhw`.
6. Also predict the standard error.
7. `predict()` only returns the predicted values. For plotting, add again the `x` values stored in `ndat_coralcover_dhw`. Now you are set up for the plot.
8. Store in `aes()` the axes that will be used in all layers. `max_dhw` should be plotted on the x axes.
9. Add a `geom_ribbon()` for the predicted data ± the standard error (`fit - se.fit` for `ymin`and `fit + se.fit` for `ymax`). These columns are in `pred_coralcover_dhw`.
10. Now, add the regression line with `geom_line()`. The column `fit` in `pred_coralcover_dhw` should be used as y axis.
11. Then, add the raw data that is stored in `dat_change_coral_cover`. Use `rel_change` as y axis.
```


:::
:::::



